p-NLI: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-21 paper/intro.tex
update intro

% (What is the high level motivation)
% What is the obsatcle for effectively apply machine learning model
With the recent advances in neural network based model, machine learning has
gained unprecedented popularity, and its adaption permeated many fields of studies.
%
However, from researchers to practitioners, one often need to overcome many
obstacles during the training, debugging, and tuning processes to unleash
the full potential of these models.
%
Identify where and how does the error occur and come up with hypothesis for
why does the model fail is critical to the design and application of various models.
%quickly identify where the errors are
However, providing meaningful answers to these questions is an extremely challenging task.
% Naturally the identification of failure scenarios is an essential first step in diagnosing procedure.
%the potential limitation of the test dataset
On the one hand, recognize where the model fail is non trivial. The standard evaluation approach provides limited information. i.e., the performance number (e.g., prediction accuracy on the test set) alone cannot answer many important questions in the model diagnostic process, such as what the often made mistakes are, and how sensitive are the model concerning small perturbation in the input, etc. On the other hand, infer why the model fails is challenging as well (i.e., models are often referred to as black boxes),
which require the ability to peek inside the model and understand the relationship between the
model input, critical internal mechanisms, as well the output predictions.

%%%%%%%%% accessing internal states are important %%%%%%%%%%%%
The wide application of machine learning models also call for increasing demands for model accountability (e.g., what is the evidence for making the decision) and model fairness (e.g., is the prediction affected by the bias in the training data).
% Making sense and explaining predictions made by neural networks is also becoming a necessity.
%
Many works~\cite{RibeiroSinghGuestrin2016} have been dedicated to providing intuitive explanations for a given prediction, several of which approach the problem from a model agnostic approach that made them applicable to different applications (i.e., by fitting a simplier linear model near the prediction of interests).
%
Despite being invaluable for the providing human understandable explaination, the lack of the ability to access and explore the internal states of the model, which is vitally important to make sense of why a model fail, limits their ability for more in depth exploration and evaluation of the model.

%%%%%%%%%% interrogate relationship between different componenet, modified different componenets %%%%%%%%%%%
Beside looking into the internal states, understanding how each of the component of the model interact with input and output as well as with each other is the crucial for truly exmaine the mechanism of the model.
%
Look at how the model work in action and interrogate relationship between different component, i.e., how the change made to one part of model affect the other pieces present, is the key to gain the full picture.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{Language Inference Model Pipeline.}
\label{fig:projTransition}
\vspace{-5mm}
\end{figure}

>>>>

% (What is the high level motivation)
% What is the obsatcle for effectively apply machine learning model
With the recent advances in neural network based model, machine learning has
gained unprecedented popularity, and its adaption permeated many fields of studies.
%
However, from researchers to practitioners, one often need to overcome many
obstacles during the training, debugging, and tuning processes to unleash
the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical to both the design and deployment of a given model.
More importantly, the ability to pinpoint where and how an error is made and come up with hypothesis for the cause of failure is the key to identify the limitation and improve upon existing models.
However, providing meaningful answers to these questions are extremely challenging and often regarded as impossible by many.

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{} on interpretation of convolution neural network (CNN) model from both machine learning and visualization community has reveal the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know regarding the inner mechanism of these deep neural network models. 
%
Many of the existing techniques study the model as an invariant object, where the model's behavior are recorded and analyzed in an offline fashion.
%
However, the exploration natural of these type of analysis often lead to \emph{what if ...} types of questions. What if I perturbed the current input? Will the prediction be stable? What if when change the internal state of the model, how could the change affect the prediction. What if the current prediction is wrong, how and where should we change the model in order to produce the correct result? These type of dynamic queries is a natural way for us to gain understanding and come up with hypothesis of the model mechanisms by \emph{interrogating} the relationship between the input, internal states, and output prediction. Yet we do not often need to setup extra specialized experiment to answer.

Compare to image classification task, natural language processing systems have a number of innate restriction (e.g., the discrete nature of words) that makes interpreting them even more challenging. For example, \emph{feature visualization}~\cite{}, an often deployed technique for illustrating what kind of image feature a given part of the network (e.g., neuron, layer, or channel) captures, can not be easily generalized and applied for natural language due to the requirement of a continuous solution space (i.e., the image pixel values). Even though we can encode a word as an vector~\cite{}, the word embedding space is still a discretely defined space, where the interpolation between two vectors (i.e., two words) do not corresponds to explicit meaning.

On the one hand, recognize where the model fail is non-trivial. The standard evaluation approach (i.e., the prediction accuracy on the test set) provides limited information and cannot answer many important questions (e.g., how stable the prediction is, which part of the model may contribute to the failure) that are essential to the model diagnostic process. On the other hand, infer the cause of model failure is even more challenging. Neural network models are often referred to as black boxes due to the hard to recognize  difficult to interpret parameter, hypothesizing what happened in prediction process require the ability to peek inside the model and understand the intricate relationships among model inputs, critical internal mechanisms, as well the output predictions.

%\begin{itemize}
%    \item Study the classifier  vs. single instance evaluation
%    \item Model agnostic vs. reveal internal states
%    \item View model as invariant object vs. interrogate the model via partial/full updates
%\end{itemize}

Questions such as, how stable predictions are concerning small perturbation in the input, how are these predictions arrived (i.e., the model may come up with the correct prediction by pickup unintended feature/alignment), and what are the often made mistakes, are all essential in systematic examination and evaluation of the model. 
%%%%%%%%% accessing internal states are important %%%%%%%%%%%%

% Making sense and explaining predictions made by neural networks is also becoming a necessity.
%
Many works~\cite{RibeiroSinghGuestrin2016} have been dedicated to providing intuitive explanations for a given prediction, several of which approach the problem from a model agnostic approach that made them applicable to different applications (i.e., by fitting a simplier linear model near the prediction of interests).
%
Despite being invaluable for the providing human understandable explaination, the lack of the ability to access and explore the internal states of the model, which is vitally important to make sense of why a model fail, limits their ability for more in depth exploration and evaluation of the model.

%%%%%%%%%% interrogate relationship between different components, modified different components %%%%%%%%%%%
Beside looking into the internal states, understanding how each of the component of the model interact with input and output as well as with each other is the crucial for truly exmaine the mechanism of the model.
%
Look at how the model work in action and interrogate relationship between different component, i.e., how the change made to one part of model affect the other pieces present, is the key to gain the full picture.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{Language Inference Model Pipeline.}
\label{fig:projTransition}
\end{figure}

### 2018-03-21 paper/relatedWorks.tex
update intro

\section{Related Works}

@@

Depth learning vis
\begin{itemize}
    \item ActiVis: Visual Exploration of Industry-Scale Deep Neural Network Models
    \item Analyzing the Training Processes of Deep Generative Models
    \item Visual Diagnosis of Tree Boosting Methods
    \item TreePOD: Sensitivity-Aware Selection of Pareto-Optimal Decision Trees
    \item LSTMVis: A Tool for Visual Analysis of Hidden State Dynamics in Recurrent Neural Networks
    \item Do Convolutional Neural Networks learn Class Hierarchy?
    \item DeepEyes: Progressive Visual Analytics for Designing Deep Neural Networks
\end{itemize}

>>>>

\section{Related Works}
Due to the significant implication of model interpretability, a wealth of researches of be focused on the explanation and interpretation of deep neural network models.

%%%% summarize analyzing prediction, but not sensitivity %%%%
Several previous works~\cite{BilalJourablooYe2018} have dedicate to analyzing errors made in analyzing

With the increasing demands for model accountability (e.g., what is the evidence for making the decision) and model fairness (e.g., is the prediction affected by the bias in the training data), these techniques only provide us a glimpse of what to come.

%%%% summarize analyzing prediction, but not sensitivity %%%%
LSTMvis~\cite{StrobeltGehrmannHuber2016} making sense of internal states of the model

There has been a
To analyzing the
Many method has been

%%%% interactive steering of the model %%%%

\subsection{Model Interpretability}

@@

\subsection{Visualization and HCI}
Depth learning vis
\begin{itemize}
    \item ActiVis: Visual Exploration of Industry-Scale Deep Neural Network Models
    \item Analyzing the Training Processes of Deep Generative Models
    \item Visual Diagnosis of Tree Boosting Methods
    \item TreePOD: Sensitivity-Aware Selection of Pareto-Optimal Decision Trees
    \item LSTMVis: A Tool for Visual Analysis of Hidden State Dynamics in Recurrent Neural Networks
    \item Do Convolutional Neural Networks learn Class Hierarchy?
    \item DeepEyes: Progressive Visual Analytics for Designing Deep Neural Networks
\end{itemize}
